---
title: "Graphics with ggplot2"
author: "Javier Beltran"
date: "18 de noviembre de 2018"
output: html_document
---

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Learning ggplot2

We're learning ggplot2 It's going to be amazing. 

# Loading data

```{r}
#National Parks in California
ca <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/ca.csv") 

#Acadia National Park
acadia <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/acadia.csv")

#Southeast US National Parks
se <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/se.csv")

#2016 Visitation for all Pacific West National Parks
visit_16 <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/visit_16.csv")

#All Nationally designated sites in Massachusetts
mass <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/mass.csv")
```

# Readin the data

```{r}
head(ca)
```

#Our first ggplot graph

Basic map with two "continuous variables":
```{r}
ggplot(data=ca) +
  geom_point(aes(x = year, y = visitors))

#We get exactly the same result writing
ggplot(ca) +
  geom_point(aes(year,visitors))

```

Distinguishing by "factor" _park_ using the "color" argument:

```{r}
ggplot(data=ca) +
  geom_point(aes(x = year, y = visitors, color = park_name))
```

Improvements

```{r}
ggplot(data=ca) +
  geom_point(aes(x = year, y = visitors, color = park_name)) +
  labs(x = "Year",
       y = "Visitation",
       title = "California National Visitation") +
  theme_bw() +
  theme(legend.title = element_blank())
```
## Exercise (10 min)
1. Using the se dataset, make a scatterplot showing visitation to all national parks in the Southeast region with color identifying individual parks.
2. Change the plot so that color indicates state.
3. Customize by adding your own title and theme. You can also change the text sizes and angles. Try applying a 45 degree angle to the x-axis. Use your cheatsheet!
4. In the code below, why isn’t the data showing up?

```{r}
# 1 
ggplot(data=se) +
  geom_point(aes(x = year, y = visitors, color = park_name))

# 2 
ggplot(data=se) +
  geom_point(aes(x = year, y = visitors, color = state))

# 3
ggplot(data=se) +
  geom_point(aes(x = year, y = visitors, color = state)) +
  labs(x = "Year",
       y = "Visitation",
       title = "Southeast Visitation") +
  theme_light() +
  theme(legend.title = element_blank(),
        axis.title.x = element_text(angle = 45, hjust = 1, size = 14))
```

# Faceting

```{r}
ggplot(data = se) +
    geom_point(aes(x = year, y = visitors)) +
    facet_wrap(~ state)
```

```{r}
ggplot(data = se) +
     geom_point(aes(x = year, y = visitors, color = park_name)) +
     facet_wrap(~ state, scales = "free")
```

# Other geoms

```{r}
ggplot(data = se) + 
  geom_jitter(aes(x = park_name, y = visitors, color = park_name), 
              width = 0.1, 
              alpha = 0.4) +
  coord_flip() +
  theme(legend.position = "none") 
```

```{r}
ggplot(se, aes(x = park_name, y = visitors)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

To visualize data over time

```{r}
ggplot(se, aes(x = year, y = visitors, color = park_name)) +
   geom_line()
```

Adding multiple geoms

```{r}
ggplot(data = acadia) + 
  geom_point(aes(x = year, y = visitors)) +
  geom_line(aes(x = year, y = visitors)) +
  geom_smooth(aes(x = year, y = visitors)) +
  labs(title = "Acadia National Park Visitation",
       y = "Visitation",
       x = "Year") +
  theme_bw()
```

```{r}
ggplot(data = acadia, aes(x = year, y = visitors)) + 
  geom_point() +
  geom_line() +
  geom_smooth(color = "red") +
  labs(title = "Acadia National Park Visitation",
       y = "Visitation",
       x = "Year") +
  theme_bw()
```

Bar charts

```{r}
ggplot(data = visit_16, aes(x = state)) + 
  geom_bar()
```

Filling the bars:

```{r}
ggplot(data = visit_16, aes(x = state, y = visitors, fill = park_name)) + 
  geom_bar(stat = "identity")
```

By default, we get a "stacked bar plot". If we want bars next to the other:

```{r}
ggplot(data = visit_16, aes(x = state, y = visitors, fill = park_name)) + 
  geom_bar(stat = "identity", position = "dodge")
```

Using mass data set and other options

```{r}
#how many of each types of parks are in Massachusetts?
ggplot(data = mass) + 
      geom_bar(aes(x = type, fill = park_name)) +
  labs(x = "",
       y = "")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7))
```

# Arranging and exporting plots

After creating your plot, you can save it to a file in your favorite format. The Export tab in the Plot pane in RStudio will save your plots at low resolution, which will not be accepted by many journals and will not scale well for posters.

Instead, use the ggsave() function, which allows you easily change the dimension and resolution of your plot by adjusting the appropriate arguments (width, height and dpi):

```{r}
my_plot <- ggplot(data = mass) + 
      geom_bar(aes(x = type, fill = park_name)) +
  labs(x = "",
       y = "")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7))

ggsave("figures/name_of_file.png", my_plot, width = 15, height = 10)

#Note: The parameters width and height also determine the font size in the saved plot.
```

#Interactive plots

```{r}
#install.packages("plotly")
library(plotly)

ggplotly(my_plot)
```

```{r}
acad_vis <- ggplot(data = acadia, aes(x = year, y = visitors)) + 
  geom_point() +
  geom_line() +
  geom_smooth(color = "red") +
  labs(title = "Acadia National Park Visitation",
       y = "Visitation",
       x = "Year") +
  theme_bw()

ggplotly(acad_vis)
```

